Ab Initio Effective One-Electron Potentials Revisited: General Theory and Example Applications for Charge-Transfer Energy Calculations

### 2019-05-22 papers/2_oep_technique/paper-r0.tex
More about OEP definition

The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
under the localized basis set approximation 
from Eq.~\eqref{e:delta-p-part}.
%Coulomb\hyp{}like and overlap\hyp{}like 
%terms from Eq.~\eqref{e:delta-p-part-2el}. 
For example, the Coulomb\hyp{}like contributions
can be rewritten without making any further approximation as
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
 \equiv
	\sum_{kl\in B} 
	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
\end{equation}
%
Similarly for the overlap\hyp{}like term we have
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
 \equiv
	\sum_{k\in A} \sum_{l\in B} 
	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
and the associated effective potential reads
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
\end{equation}
%
Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
%
\begin{multline}
 \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
	+ \sum_{i\in A} \sum_{j\in B} 
	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
	 \\ +
	\sum_{i\in A} \sum_{k\in A}
	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
	 +
        \sum_{j\in B} \sum_{l\in B}
          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
\end{multline}
%
where the exchange\hyp{}like component is
%
\begin{equation}
	\Delta_{\rm ex}^{AB} P = 
	s \sum_{ik\in A} \sum_{jl\in B}
	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
\end{equation}
%
and
the overlap\hyp{}like and the Coulomb\hyp{}like 
OEP's are defined as follows
%
\begin{subequations}
\begin{align}
	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \;, \\
	%
	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
        \\
	&+ \frac{1}{2}
        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
\end{align}
\end{subequations}
%
and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.

@@

There are two general cases of basis function partitioning scheme
in which it is possible to define an effective one\hyp{}electron
potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, can be approximated with the density fitting. Mainly,
%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{{\rm eff},A}}{j} $ && $\tBraKet{j}{\hat{v}^{{\rm eff},A}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &   \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%

>>>>

The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
under the localized basis set approximation 
from Eq.~\eqref{e:delta-p-part}.
%Coulomb\hyp{}like and overlap\hyp{}like 
%terms from Eq.~\eqref{e:delta-p-part-2el}. 
For example, the Coulomb\hyp{}like contributions
can be rewritten without making any further approximation as
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
 \equiv
	\sum_{kl\in B} 
	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
\end{equation}
%
Similarly for the overlap\hyp{}like term we have
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
 \equiv
	\sum_{k\in A} \sum_{l\in B} 
	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
and the associated effective potential reads
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
\end{equation}
%
Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
%
\begin{multline} \label{e:ci-oep-master}
 \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
	+ \sum_{i\in A} \sum_{j\in B} 
	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
	 \\ +
	\sum_{i\in A} \sum_{k\in A}
	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
	 +
        \sum_{j\in B} \sum_{l\in B}
          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
\end{multline}
%
where the exchange\hyp{}like component is
%
\begin{equation}
	\Delta_{\rm ex}^{AB} P = 
	s \sum_{ik\in A} \sum_{jl\in B}
	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
\end{equation}
%
and
the overlap\hyp{}like and the Coulomb\hyp{}like 
OEP's are defined as follows
%
\begin{subequations}
\begin{align} \label{e:oep-exch-like}
	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \;, \\
	%
	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
        \\ \label{e:oep-coul-like}
	& \qquad\qquad+ \frac{1}{2}
        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} \;,
\end{align}
\end{subequations}
%
and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.

\subsection{\label{s.242}Separability of P-Coefficients}

Until that point, the effective potentials from Eqs.~\eqref{e:oep-exch-like}
and \eqref{e:oep-coul-like} are not separable in terms of the interacting molecules
because the $\bf P$ tensors are derived from the CI wavefunction of
the entire interacting system $AB$. However, if there exist an approximate
separation as, for example,
%
\begin{align}
 P^{AAAB}_{mkij} &\approx X^A_{mk} Y^{AB}_{ij} \;, \\
 P^{AABB}_{ikjl} &\approx Z^A_{ik} Y^{AB}_{} Z^{ B}_{jl} \;,
\end{align}
%
and if the matrices ${\bf X}^{A(B)}$, 
${\bf Z}^{A(B)}$ and ${\bf P}^{AA(BB)}$ as well as the tensors
${\bf P}^{AAAA(BBBB)}$ 
are properties of the unperturbed (isolated) molecules,
further rearrangements in Eq.~\eqref{e:ci-oep-master} are possible
leading to computationally more efficient expressions.

Note also that the matrices ${\bf Y}^{AB}$ and ${\bf P}^{AB}$ depend on their instantaneous 
relative positions.

%then the molecule\hyp{}specific OEP's can be defined.

@@

There are two general cases of basis function partitioning scheme
in which it is possible to define an effective one\hyp{}electron
potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. Mainly,
%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{{\rm eff},A}}{j} $ && $\tBraKet{j}{\hat{v}^{{\rm eff},A}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &   \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%
